\textsf{\textbf{Drop Electro-Bouncing in Low-Gravity}}

### 2018-04-05 doc/manuscript/manuscript.tex
many minor edits its to manuscript

\section*{Abstract}
\addcontentsline{toc}{chapter}{\numberline{Abstract}}
\noindent
We investigate the dynamics of spontaneous jumps of water drops from electrically charged superhydrophobic dielectric substrates during a sudden step reduction in gravity level. In the brief free-fall environment of a drop tower, with a non-homogeneous external electric field arising due to dielectric surface charges (with surface potentials ~0.4-1.8 kV), body forces acting on the jumped drops are primarily supplied by polarization stress and Coulombic attraction instead of gravity. This electric body force leads to a drop bouncing behavior similar to well-known phenomena in 1-$g_0$, though occurring for much larger drops ($\sim$ 0.5 mL). We show a simple model for the phenomenon, its scaling, and asymptotic estimates for drop time of flight in two regimes: at short-times close to the substrate when drop inertia balances Coulombic force due to net free charge and image charges in the dielectric substrate and at long-times far from the substrate when drop inertia balances free charge Coulombic force and drag. The drop trajectories are controlled primarily by the dimensionless electrostatic Euler number $\mathbb{E}\mbox{u}$, which is a ratio of inertial to electrostatic forces. To experimentally determine values of $\mathbb{E}\mbox{u}$ we conduct a series of drop tower experiments where we observe the effects of drop volume, net free charge, and static surface potential of the superhydrophobic substrate on drop trajectories. We use a direct search optimization to obtain a Maximum Likelihood Estimate for drop net charge, as we do not measure it directly in experiment. For $\phi \mathbb{E}\mbox{u} /8 \pi > 1$ drops escape the electric field, where $\phi$ is a drop to substrate aspect ratio. However, we do not observe any escapes in our dataset. With an eye towards engineering applications we consider the results in light of the so-called low-gravity phase separation problem with a worked example problem.

@@

As with capillary forces, electrostatic forces have a multitude of applications as substitutes for gravity. Electrohydrodynamic (EHD) body forces have been suggested to combat thermal stratification of propellants under low-g conditions where bouyancy induced natural convection is very small \cite{blackmon_collection_1965}. In this case electro-convection is an analog of natural convection due to a fluid temperature dependent dielectric permittivity gradient (as opposed to a fluid density gradient) in the presence of an electric body force field. The study also suggested applications in cryogenic tank vent screening and in reducing heat transfer to cryogenic propellants (and thus boiloff losses) by dielectrophoretically positioning ullage vapors around the tank walls as a thermal barrier. EHD forces have been studied for enhancing boiling heat transfer by promotion of bubble detachment and prevention of dryout, again as an ersatz bouyancy \cite{snyder_dielectrophoresis_2001, di_marco_influence_2003, marco_use_2012}. EHD heat pipes, which substitute an electrode structure for the capillary wicking structure of a conventional thermocapillary heat pipe can postpone the wicking limit \cite{jones_electrohydrodynamic_1973}. These are restricted to the use of insulating dielectric liquids, which usually have relatively low thermal conductivity, but are highly robust against dryout, support active flow control, and have low viscous losses. Dielectrophoretic settling of cryogenic propellants, in both total and partial communication Propellant Management Device (PMD) configurations, has been studied analytically in both static \cite{hurwitz_electrohydrodynamic_1966} and dynamic cases \cite{koval_dynamics_1967}, and in drop tower experiments \cite{fax_dielectrophoretic_1969}. EHD forces have also been applied to the low-g slosh baffling problem \cite{boretz_orbital_1970, petrash_use_1968, hurwitz_dielectrophoretic_1968}, and studied extensively for the mitigation of vapor pullthrough during tank draining and concomittent minimization of propellant residuals at burnout in a series of experiments at the NASA Lewis (now Glenn) Research Center 2.2 s drop tower facillity \cite{berenyi_dielectrophoretic_1970}.

\section{Spontaneous Drop Jump}
Low-gravity phase separation, drop dynamics, and the dynamics and applications of the so-called spontaneous drop or puddle jump, is an active area of study at the Portland State University Dryden Drop Tower labratory (DDT). When a nonwetting, gravity-dominated sessile drop (e.g. a puddle), which is initially at rest on a surface in the Cassie-Baxter state suddenly undergoes a large step reduction in $\mathbb{B}\mbox{o}$ it will spontaneously jump away from the surface. The Cassie-Baxter state, sometimes also known as the Fakir state, is a metastable heterogenously non-wetting state on a textured surface characterized by the existence of a vapor-layer that seperates roughness features of the surface from the bottom of the drop. The spontaneous drop jump was first observed experimentally in the Soviet Union by Kirko \emph{et al.} \cite{kirko_phenomenon_1970} in 1970 for drops of mercury in hydrochloric acid, and later by Wollman \emph{et al.} in 2016 for water drops in air in a set of experiments conducted using drop towers \cite{wollman_more_2016}. The kinetic energy of the jump is supplied by the defect in free surface energy as the new minimum energy surface equilibrium has approximately constant curvature. If the energy losses by viscous and contact line dynamics are neglected, the available kinetic energy is approximated by
\[KE_2 = SE_2 - SE_1 = [(\gamma A)_{gl} + (\gamma A)_{sg}]_2 - [(\gamma A)_{ls} + (\gamma A)_{gl} + (\gamma A)_{sg}]_1, \]
where $KE$ and $SE$ are the respective kinetic and surface energies, $A$ is the surface area, the subscripts $1$ and $2$ denote the intial 1-$g_0$, and low-g states, and the subscripts ${gl}$, ${sg}$, and ${ls}$ denote the gas-liquid, solid-gas and liquid-solid interfaces repectively. The motive force is due to the inertia of internal flows which occur as the interface deflects under the suddenely lessened $\mathbb{B}\mbox{o}$. The characteristic time scale of the rolling up scales as $t_j \sim (\rho V_d/\sigma)^{1/2}$ \cite{attari_puddle_2016}, which resembles the contact time, $\tau \approx 2.6(\rho R^3_d/\sigma)^{1/2}$ presented by Richard \emph{et al.} in 2002 for the related problem of drops impacting hydrophobic surfaces in 1-$g_0$ \cite{richard_surface_2002}. For drops with radial symmetry and sufficiently high initial $\mathbb{B}\mbox{o}$, inward radial capillary waves coalesce at the axis, leading to geysering and creation of satellite droplets by the Rayleigh-Plateau breakup of the geyser. In the case of smaller jumping drops the capillary waves are gradually damped by viscous forces. A time-lapsed composite photograph of spontaneous drop jump can be seen in Figure \ref{fig:nobounce}.
\begin{figure}[htb]
\centering
\includegraphics[width=0.75\textwidth]{drop_jump}
\caption{The trajectory of a 0.2 mL spontaneous drop jump is captured in a composite image over a period $\sim 0.67$ s presented at $\approx 10$ Hz. \label{fig:nobounce}}
\end{figure}

@@

There are several possible origins of the electric charge that are ultimately responsible for this phenonenon. It is well known that water acquires positive free charge when in contact with certain polymers, especially polytetrafluoroethylene (PTFE) \cite{langmuir_surface_1938}, through a process called contact charging. PTFE, on the other hand, tends to readily acquire negative charge by contact with water. The superhydrophobic surfaces used in the spontanous drop jump experiments have thin (nanometric) PTFE coatings, and we observe that it is extremely easy to produce significant surface potentials $\varphi_s \sim$ 100-500 V by simply flowing streams of distilled water over them. A study of this water on PTFE contact charging phenomenon was conducted by Yatsuzuka \emph{et al.} \cite{yatsuzuka_electrification_1994}, who suggest that this process results from formation of an electrical double layer driven by selective adsorption of ($\mbox{OH}^-$) ions at the polymer surface; other recent work strongly supports this hypothesis \cite{beattie_intrinsic_2006, strazdaite_water_2015}. Yatsuzuka \emph{et al.} also found that the specific charge on the drops in contact with the surface depends on their both surface velocity and conductivity. The contact charging mechanism is the most likely accidental source of charge on the suphydrophobic surfaces used in our work. Given the large roughness, or the ratio of projected to actual surface area, of the superhydrophobic surfaces used in the experiment, and given that the drops are initially in a Cassie-Baxter state, a somewhat electrically resistive air layer is maintained that reduces grounding of the drops despite the large potential difference between them and the surface charges. If this electrostatic potential is greater than the release of surface energy under the sudden change in $\mathbb{B}\mbox{o}$, in the form of the drop kinetic energy, we expect that the drop will return to an equilibrium state on the charged surface which minimizes the potential.

The source of the net free charge on the drops is another issue. The drop charge could also be due to the contact charging mechanism mentioned previously. For instance, in a 1996 paper, NASA flight engineer Don Pettit discusses the problem of low-gravity flow induced charging of liquids, resulting ultimately from contact charging phenomenon \cite{pettit_donald_flow_????}. Another mechanism for the drop charge is field-induced charging. Field-induced charging occurs due to physical breakup of a conductor having a field-induced dipole (e.g., a physical seperation of charge). In our work this might occur when a drop is deposited on the charged surface by a grounded syringe. The metal syringe needle tip, and the liquid in the syringe itself are essentially a ground connection which is broken when the syringe is suddenly removed in the presence of an external electric field. Field-induced charging is at work in the famous Kelvin thunderstorm, and is applied in inkjet, and electrospay technologies, where in each case the breakup is by the Rayleigh-Plateau instability. Notably, in Pettit's aforementioned discussion of contact charging of liquids in low-g, he remarks on accidental electrostatic `hula-ing' of silicone oil drops when ejected from a syringe in the vicinity of a highly-charged polymer surface during an experiment conducted aboard STS-5 by mission specialist Joseph Allen \cite{pettit_donald_flow_????}. Depending on the (highly-variable) electrical conductivity of the silicone oil, and the material of the syringe used in the experiment, the charge could arise just as easily by field-induction as by contact charging. Relatedly, in a series of informal and somewhat whimsical experiments Pettit himself electrostatically orbited small water drops around a triboelectrically charged PTFE knitting needle while aboard ISS during expedition 30/31 \cite{stevenson_electrostatic_2015}. Again, in this case, the drop charge is likely field-induced.

\section{Applications}
This study of the electro-drop bounce phenomenon may have results which can be extended to useful applications in a space environment. External surfaces of spacecraft tend to become charged with time due to the space environment itself. Surface charging largely occurs due to low energy electrons (3-50 keV) which do not penetrate the surface of the spacecraft external stucture \cite{czepiela_charging_1997}. The ultimate sources of these charges are trapped charged particles of the van Allen radiation belts, galactic cosmic rays, and the solar wind. This deposited charge accumulates and can lead to significant potential differences between different parts of a spacecraft, sometimes leading to breakdown and discharge called Paschen discharge. Deep dielectric charging occurs when higher energy charged particles penetrate the surface of a dielectric material, which can also lead to large potential differences if the dielectric leakage is lower than the external charging rate. In a fluid mechanical context, this accumulated charge could be problematic during any multiphase venting process, as might be expected to occur during autonomous satellite servicing refueling operations, or during boil-off venting of cryogens stored over long periods in a space environment (as in many crewed Mars mission scenarios, or in cryogenic propellant depot operations). Such venting, if occurring in a region of strong field, will also tend to produce a stream of charged drops, which will be decelerated by the field, causing their eventual return, impacting and potentially contaminating the spacecraft surfaces.

As previously noted, robust phase separation is critical to high reliability multiphase systems used in thermal control and life support. Active solid-state phase seperation for life support multiphase flows, especially high void-fraction droplet disperse flows is a possible application of this work. Such flows are encountered in dish washing, laundry, waste solids drying, food processing, the Sabatier CO$_2$ reaction, and possibly in vapor-compression cycle condensers. Phase seperation for other disperse droplet flows include electrostatic droplet seperators for high-efficiency Rankine-cycle turbines, which require a droplet-free vapor phase entering the turbine, but where conventional superheat approaches come with severe mass penalties \cite{unterberg_zero_1962}. More speculatively, results of this work could be applied in high temperature liquid droplet radiators with electrostatic collection \cite{white_liquid_1987}. Removal of satellite droplets produced during pipetting in wet-lab research outside of a glovebox environment aboard ISS has also been recently suggested as a application of the work [private communication, NASA JSC, E. Unger, 2017]. Drops can become spontaneously charged by contact with standard micropipette tips \cite{choi_spontaneous_2013}, and this free charge can possibly be leveraged for the purposes of phase separation in low-gravity.

@@

It is common to approximate the polarization stress by idealizing the drop as a simple dipole using the effective dipole moment method first described by Pohl in 1958 \cite{pohl_effects_1958}. This approach has since been related back to the force density by the volumetric integration of the force density with the substitution of a Taylor series expansion of $\mathbf{E}$ in the limit of a small radial gradient of the field within the dielectric sphere \cite{wang_general_1997}. The DEP force is related to the induced or permanent dipole moment of polarizable media which has a tendency to align the dipole with the electric field. If there is a gradient in the field then for a finite separation of charge, one end of the dipole will feel a stronger electric field than the other resulting in a net force. Whether the force is positive or negative in the direction of the electric field gradient depends on the difference of dielectric permittivites between the fluids, rather than on the polarity of $\mathbf{E}$. In principle an external electric field will tend to induce a dipole in a dielectric material, but if the field is spatially uniform there is no gradient in the field and the forces felt by the dipoles are symmetric with no net force. The dipole moment $\mu$ of a spherical linear-dielectric particle immersed in a dielectric medium is given by
\begin{equation} \label{dipole_m_1}
\mu = V_d \mathbf{P} = \frac{4}{3} \pi R_d^3 \mathbf{P},
\end{equation} 
where $\mathbf{P} = \left(\kappa_1 - 1 \right) \epsilon_0 \mathbf{E} = \chi_e \epsilon_0 \mathbf{E}$ is the polarization moment, $R_d$ is the particle radius, $\kappa_1 = \frac{\epsilon_1}{\epsilon_0}$ is the relative dielectric constant of the air host fluid, and $\chi_e = \kappa_1 - 1$ is the electric susceptibility of the host fluid. The force felt by the dipole is 
\begin{eqnarray} \label{dep_force}
\mathbf{F}_{DEP} &=& \left( \mathbf{P}_e \cdot \nabla \right) \mathbf{E} \nonumber \\
&=& 2 \pi R_d^3 \epsilon K \nabla E^2,
\end{eqnarray}
where $\mathbf{P}_e=(\kappa_2 - \kappa_1)\epsilon_0 \mathbf{E}$ is the excess polarization and $\kappa_2$ is the relative dielectric constant of the water particle. Here it is convenient to use the simplifying shorthand $K = \frac{\kappa_2 - \kappa_1}{\kappa_2 + 2 \kappa_1}$, known as the Clausius-Mossotti factor. In cases where $K <$ 0, or $K>$ 0 the particle will be either repelled or attracted to regions of strong field. In our experiment, choosing the relative dielectric constants $\kappa_1 \approx$ 1 and $\kappa_2 \approx$ 80, we estimate $K \approx$ 0.96. It is important to note that the equivalent dipole approximation critically requires an assumption of small physical scale of the particle relative to the length scale of nonuniformity of the field, which in this case we take to be the length of the charged superhydrophobic surface, $L =$ 25 mm $\gg R_d \approx$ 2 mm.

When the drop is close to the dielectric surface, the free charge on the drop will tend to induce polarization of the dielectric which perturbs the electric field. The polarization bound charge in the dielectric will be of the opposite sign of the free drop charge and thus there will be a force of attraction. This so-called image force is a correction to the Coulomb force due to the external electric field only, and can be found by a Green's Function solution to Laplace's equation for the electric field called the `method of images' \cite{david_j._griffiths_introduction_1999}. This resulting image force $\mathbf{F}_I$ is given by
\begin{equation}
\mathbf{F}_I = \frac{k q^2}{16 \pi \epsilon_0} y^{-2} \hat{\mathbf{j}},
\label{image_force}
\end{equation}
where the factor $k$ is a function of the dielectric surface susceptibility $k = \frac{\chi_e}{\chi_e + 2}$, and $\hat{\mathbf{j}}$ is a unit vector normal to the dielectric surface.

By substituting Equations \ref{dep_force}, and \ref{image_force} into \ref{force_density} we have
\begin{eqnarray*}
 \mathbf{F}_E &=& q \mathbf{E} + \mathbf{F}_{DEP} + \mathbf{F}_I \\
 &=& q \mathbf{E} + \frac{k q^2}{16 \pi \epsilon_0 } y^{-2} \hat{\mathbf{j}} + 2 \pi R_d^3 \kappa_1 \epsilon_0 K \nabla E^2, 
\end{eqnarray*}
and the 1-D governing equation becomes
\begin{equation}
 \label{gov_eqn_subs}
m y'' = - \frac{1}{2} C_D \rho A {y'}^2 - q E - \frac{k q^2}{16 \pi \epsilon_0} y^{-2}- 2 \pi R_d^3 \kappa_1 \epsilon_0 K \nabla E^2,
\end{equation}
subject to
\begin{equation}
y(0) = R, \hspace{5 mm} \mbox{and} \hspace{5 mm} y'(0) = U_0 .
\end{equation}

By comparing DEP and Coulombic terms in Equation \ref{gov_eqn_subs}, we note that a condition to neglect the DEP force is
\begin{eqnarray}
\frac{ \kappa_2 \epsilon_0 K R_d^2 E_0}{q} \ll 1. \nonumber
\end{eqnarray}
As this condition likely prevails in our experiments we henceforth neglect the DEP force. There is some physical intuition to support this conclusion as well. The dielectric displacement $\mathbf{D} = \kappa \epsilon_0 \mathbf{E}$ of a water drop in air is high due to its large relative dielectric constant. This implies that the field strength within the drop is about 80 times smaller than in the surrounding medium, which is essentially the same as a vacuum from a dielectric standpoint. Thus it is not particularly inaccurate to model the dielectric volume of a drop as an equipotential conductive shell with zero field in its interior. As an aside, in treating the drop as an ideal conductor we note that in our specific case the electrostatic force is not a body force \emph{per se} as the electric field is acting on charges on the surface of the conductor.

>>>>

\section*{Abstract}
\addcontentsline{toc}{chapter}{\numberline{Abstract}}
\noindent
We investigate the dynamics of spontaneous jumps of water drops from electrically charged superhydrophobic dielectric substrates during a sudden step reduction in gravity level. In the brief free-fall environment of a drop tower, with a non-homogeneous external electric field arising due to dielectric surface charges (with surface potentials ~0.4-1.8 kV), body forces acting on the jumped drops are primarily supplied by polarization stress and Coulombic attraction instead of gravity. This electric body force leads to a drop bouncing behavior similar to well-known phenomena in 1-$g_0$, though occurring for much larger drops ($\sim$ 0.5 mL). We show a simple model for the phenomenon, its scaling, and asymptotic estimates for drop time of flight in two regimes: at short-times close to the substrate when drop inertia balances Coulombic force due to net free charge and image charges in the dielectric substrate and at long-times far from the substrate when drop inertia balances free charge Coulombic force and drag. The drop trajectories are controlled primarily by the dimensionless electrostatic Euler number $\mathbb{E}\mbox{u}$, which is a ratio of inertial to electrostatic forces. To experimentally determine values of $\mathbb{E}\mbox{u}$ we conduct a series of drop tower experiments where we observe the effects of drop volume, net free charge, and static surface potential of the superhydrophobic substrate on drop trajectories. We use a direct search optimization to obtain a Maximum Likelihood Estimate for drop net charge, as we do not measure it directly in experiment. For $\phi \mathbb{E}\mbox{u} /8 \pi > 1$ drops escape the electric field, where $\phi$ is a drop to substrate aspect ratio. However, we do not observe any escapes in our dataset. With an eye towards engineering applications we consider the results in light of the so-called low-gravity phase separation problem with a worked example.

@@

As with capillary forces, electrostatic forces have a multitude of applications as substitutes for gravity. Electrohydrodynamic (EHD) body forces have been suggested to combat thermal stratification of propellants under low-g conditions where buoyancy induced natural convection is very small \cite{blackmon_collection_1965}. In this case electro-convection is an analog of natural convection due to a fluid temperature dependent dielectric permittivity gradient (as opposed to a fluid density gradient) in the presence of an electric body force field. The study also suggested applications in cryogenic tank vent screening and in reducing heat transfer to cryogenic propellants (and thus boiloff losses) by dielectrophoretically positioning ullage vapors around the tank walls as a thermal barrier. EHD forces have been studied for enhancing boiling heat transfer by promotion of bubble detachment and prevention of dryout, again as an ersatz buoyancy \cite{snyder_dielectrophoresis_2001, di_marco_influence_2003, marco_use_2012}. EHD heat pipes, which substitute an electrode structure for the capillary wicking structure of a conventional thermocapillary heat pipe can postpone the wicking limit \cite{jones_electrohydrodynamic_1973}. These are restricted to the use of insulating dielectric liquids, which usually have relatively low thermal conductivity, but are highly robust against dryout, support active flow control, and have low viscous losses. Dielectrophoretic settling of cryogenic propellants, in both total and partial communication Propellant Management Device (PMD) configurations, has been studied analytically in both static \cite{hurwitz_electrohydrodynamic_1966} and dynamic cases \cite{koval_dynamics_1967}, and in drop tower experiments \cite{fax_dielectrophoretic_1969}. EHD forces have also been applied to the low-g slosh baffling problem \cite{boretz_orbital_1970, petrash_use_1968, hurwitz_dielectrophoretic_1968}, and studied extensively for the mitigation of vapor pullthrough during tank draining and concomittent minimization of propellant residuals at burnout in a series of experiments at the NASA Lewis (now Glenn) Research Center 2.2 s drop tower facility \cite{berenyi_dielectrophoretic_1970}.

\section{Spontaneous Drop Jump}
Low-gravity phase separation, drop dynamics, and the dynamics and applications of the so-called spontaneous drop or puddle jump, is an active area of study at the Portland State University Dryden Drop Tower laboratory (DDT). When a nonwetting, gravity-dominated sessile drop (e.g. a puddle), which is initially at rest on a surface in the Cassie-Baxter state suddenly undergoes a large step reduction in $\mathbb{B}\mbox{o}$ it will spontaneously jump away from the surface. The Cassie-Baxter state, sometimes also known as the Fakir state, is a metastable heterogenously non-wetting state on a textured surface characterized by the existence of a vapor-layer that separates roughness features of the surface from the bottom of the drop. The spontaneous drop jump was first observed experimentally in the Soviet Union by Kirko \emph{et al.} \cite{kirko_phenomenon_1970} in 1970 for drops of mercury in hydrochloric acid, and later by Wollman \emph{et al.} in 2016 for water drops in air in a set of experiments conducted using drop towers \cite{wollman_more_2016}. The kinetic energy of the jump is supplied by the defect in free surface energy as the new minimum energy surface equilibrium has approximately constant curvature. If the energy losses by viscous and contact line dynamics are neglected, the available kinetic energy is approximated by
\[KE_2 = SE_2 - SE_1 = [(\gamma A)_{gl} + (\gamma A)_{sg}]_2 - [(\gamma A)_{ls} + (\gamma A)_{gl} + (\gamma A)_{sg}]_1, \]
where $KE$ and $SE$ are the respective kinetic and surface energies, $A$ is the surface area, the subscripts $1$ and $2$ denote the intial 1-$g_0$, and low-g states, and the subscripts ${gl}$, ${sg}$, and ${ls}$ denote the gas-liquid, solid-gas and liquid-solid interfaces respectively. The motive force is due to the inertia of internal flows which occur as the interface deflects under the suddenly lessened $\mathbb{B}\mbox{o}$. The characteristic time scale of the rolling up scales as $t_j \sim (\rho V_d/\sigma)^{1/2}$ \cite{attari_puddle_2016}, which resembles the contact time, $\tau \approx 2.6(\rho R^3_d/\sigma)^{1/2}$ presented by Richard \emph{et al.} in 2002 for the related problem of drops impacting hydrophobic surfaces in 1-$g_0$ \cite{richard_surface_2002}. For drops with radial symmetry and sufficiently high initial $\mathbb{B}\mbox{o}$, inward radial capillary waves coalesce at the axis, leading to geysering and creation of satellite droplets by the Rayleigh-Plateau breakup of the geyser. In the case of smaller jumping drops the capillary waves are gradually damped by viscous forces. A time-lapsed composite photograph of spontaneous drop jump can be seen in Figure \ref{fig:nobounce}.
\begin{figure}[htb]
\centering
\includegraphics[width=0.75\textwidth]{drop_jump}
\caption{The trajectory of a 0.2 mL spontaneous drop jump is captured in a composite image over a period $\sim 0.67$ s presented at $\approx 10$ Hz. \label{fig:nobounce}}
\end{figure}

@@

There are several possible origins of the electric charge that are ultimately responsible for this phenomenon. It is well known that water acquires positive free charge when in contact with certain polymers, especially polytetrafluoroethylene (PTFE) \cite{langmuir_surface_1938}, through a process called contact charging. PTFE, on the other hand, tends to readily acquire negative charge by contact with water. The superhydrophobic surfaces used in the spontaneous drop jump experiments have thin (nanometric) PTFE coatings, and we observe that it is extremely easy to produce significant surface potentials $\varphi_s \sim$ 100-500 V by simply flowing streams of distilled water over them. A study of this water on PTFE contact charging phenomenon was conducted by Yatsuzuka \emph{et al.} \cite{yatsuzuka_electrification_1994}, who suggest that this process results from formation of an electrical double layer driven by selective adsorption of ($\mbox{OH}^-$) ions at the polymer surface; other recent work strongly supports this hypothesis \cite{beattie_intrinsic_2006, strazdaite_water_2015}. Yatsuzuka \emph{et al.} also found that the specific charge on the drops in contact with the surface depends on their both surface velocity and conductivity. The contact charging mechanism is the most likely accidental source of charge on the suphydrophobic surfaces used in our work. Given the large roughness, or the ratio of projected to actual surface area, of the superhydrophobic surfaces used in the experiment, and given that the drops are initially in a Cassie-Baxter state, a somewhat electrically resistive air layer is maintained that reduces grounding of the drops despite the large potential difference between them and the surface charges. If this electrostatic potential is greater than the release of surface energy under the sudden change in $\mathbb{B}\mbox{o}$, in the form of the drop kinetic energy, we expect that the drop will return to an equilibrium state on the charged surface which minimizes the potential.

The source of the net free charge on the drops is another issue. The drop charge could also be due to the contact charging mechanism mentioned previously. For instance, in a 1996 paper, NASA flight engineer Don Pettit discusses the problem of low-gravity flow induced charging of liquids, resulting ultimately from contact charging phenomenon \cite{pettit_donald_flow_????}. Another mechanism for the drop charge is field-induced charging. Field-induced charging occurs due to physical breakup of a conductor having a field-induced dipole (e.g., a physical separation of charge). In our work this might occur when a drop is deposited on the charged surface by a grounded syringe. The metal syringe needle tip, and the liquid in the syringe itself are essentially a ground connection which is broken when the syringe is suddenly removed in the presence of an external electric field. Field-induced charging is at work in the famous Kelvin thunderstorm, and is applied in inkjet, and electrospay technologies, where in each case the breakup is by the Rayleigh-Plateau instability. Notably, in Pettit's aforementioned discussion of contact charging of liquids in low-g, he remarks on accidental electrostatic `hula-ing' of silicone oil drops when ejected from a syringe in the vicinity of a highly-charged polymer surface during an experiment conducted aboard STS-5 by mission specialist Joseph Allen \cite{pettit_donald_flow_????}. Depending on the (highly-variable) electrical conductivity of the silicone oil, and the material of the syringe used in the experiment, the charge could arise just as easily by field-induction as by contact charging. Relatedly, in a series of informal and somewhat whimsical experiments Pettit himself electrostatically orbited small water drops around a triboelectrically charged PTFE knitting needle while aboard ISS during expedition 30/31 \cite{stevenson_electrostatic_2015}. Again, in this case, the drop charge is likely field-induced.

\section{Applications}
This study of the electro-drop bounce phenomenon may have results which can be extended to useful applications in a space environment. External surfaces of spacecraft tend to become charged with time due to the space environment itself. Surface charging largely occurs due to low energy electrons (3-50 keV) which do not penetrate the surface of the spacecraft external structure \cite{czepiela_charging_1997}. The ultimate sources of these charges are trapped charged particles of the van Allen radiation belts, galactic cosmic rays, and the solar wind. This deposited charge accumulates and can lead to significant potential differences between different parts of a spacecraft, sometimes leading to breakdown and discharge called Paschen discharge. Deep dielectric charging occurs when higher energy charged particles penetrate the surface of a dielectric material, which can also lead to large potential differences if the dielectric leakage is lower than the external charging rate. In a fluid mechanical context, this accumulated charge could be problematic during any multiphase venting process, as might be expected to occur during autonomous satellite servicing refueling operations, or during boil-off venting of cryogens stored over long periods in a space environment (as in many crewed Mars mission scenarios, or in cryogenic propellant depot operations). Such venting, if occurring in a region of strong field, will also tend to produce a stream of charged drops, which will be decelerated by the field, causing their eventual return, impacting and potentially contaminating the spacecraft surfaces.

As previously noted, robust phase separation is critical to high reliability multiphase systems used in thermal control and life support. Active solid-state phase separation for life support multiphase flows, especially high void-fraction droplet disperse flows is a possible application of this work. Such flows are encountered in dish washing, laundry, waste solids drying, food processing, the Sabatier CO$_2$ reaction, and possibly in vapor-compression cycle condensers. Phase separation for other disperse droplet flows include electrostatic droplet separators for high-efficiency Rankine-cycle turbines, which require a droplet-free vapor phase entering the turbine, but where conventional superheat approaches come with severe mass penalties \cite{unterberg_zero_1962}. More speculatively, results of this work could be applied in high temperature liquid droplet radiators with electrostatic collection \cite{white_liquid_1987}. Removal of satellite droplets produced during pipetting in wet-lab research outside of a glovebox environment aboard ISS has also been recently suggested as a application of the work [private communication, NASA JSC, E. Unger, 2017]. Drops can become spontaneously charged by contact with standard micropipette tips \cite{choi_spontaneous_2013}, and this free charge can possibly be leveraged for the purposes of phase separation in low-gravity.

@@

It is common to approximate the polarization stress by idealizing the drop as a simple dipole using the effective dipole moment method first described by Pohl in 1958 \cite{pohl_effects_1958}. This approach has since been related back to the force density by the volumetric integration of the force density with the substitution of a Taylor series expansion of $\mathbf{E}$ in the limit of a small radial gradient of the field within the dielectric sphere \cite{wang_general_1997}. The DEP force arises due to the induced or permanent dipole moment of polarizable media which has a tendency to align the dipole with the electric field. If there is a gradient in the field then for a finite separation of charge, one end of the dipole will feel a stronger electric field than the other resulting in a net force. Whether the force is positive or negative in the direction of the electric field gradient depends on the difference of dielectric permittivites between the fluids, rather than on the polarity of $\mathbf{E}$. In principle an external electric field will tend to induce a dipole in a dielectric material, but if the field is spatially uniform there is no gradient in the field and the forces felt by the dipoles are symmetric with no net force. The dipole moment $\mu$ of a spherical linear-dielectric particle immersed in a dielectric medium is given by
\begin{equation} \label{dipole_m_1}
\mu = V_d \mathbf{P} = \frac{4}{3} \pi R_d^3 \mathbf{P},
\end{equation} 
where $\mathbf{P} = \left(\kappa_a - 1 \right) \epsilon_0 \mathbf{E} = \chi_e \epsilon_0 \mathbf{E}$ is the polarization moment, $R_d$ is the particle radius, $\kappa_a = \frac{\epsilon_a}{\epsilon_0}$ is the relative dielectric constant of the air host fluid, and $\chi_e = \kappa_a - 1$ is the electric susceptibility of the host fluid. The force felt by the dipole is 
\begin{eqnarray} \label{dep_force}
\mathbf{F}_{DEP} &=& \left( \mathbf{P}_e \cdot \nabla \right) \mathbf{E} \nonumber \\
&=& 2 \pi R_d^3 \kappa_w \epsilon_0 K \nabla E^2,
\end{eqnarray}
where $\mathbf{P}_e=(\kappa_w - \kappa_a)\epsilon_0 \mathbf{E}$ is the excess polarization and $\kappa_w$ is the relative dielectric constant of the water particle. Here it is convenient to use the simplifying shorthand $K = \frac{\kappa_w - \kappa_a}{\kappa_w + 2 \kappa_a}$, known as the Clausius-Mossotti factor. In cases where $K <$ 0, or $K>$ 0 the particle will be either repelled or attracted to regions of strong field. In our experiment, choosing the relative dielectric constants $\kappa_a \approx$ 1 and $\kappa_w \approx$ 80, we estimate $K \approx$ 0.96. It is important to note that the equivalent dipole approximation critically requires an assumption of small physical scale of the particle relative to the length scale of nonuniformity of the field, which in this case we take to be the length of the charged superhydrophobic surface, $L =$ 25 mm $\gg R_d \approx$ 2 mm.

When the drop is close to the dielectric surface, the free charge on the drop will tend to induce polarization of the dielectric which perturbs the electric field. The polarization bound charge in the dielectric will be of the opposite sign of the free drop charge and thus there will be a force of attraction. This so-called image force is a correction to the Coulomb force due to the external electric field only, and can be found by a Green's Function solution of Laplace's equation for the electric field called the `method of images' \cite{david_j._griffiths_introduction_1999}. This resulting image force $\mathbf{F}_I$ is given by
\begin{equation}
\mathbf{F}_I = \frac{k q^2}{16 \pi \epsilon_0} y^{-2} \hat{\mathbf{j}},
\label{image_force}
\end{equation}
where the factor $k$ is a function of the dielectric surface susceptibility $k = \frac{\chi_e}{\chi_e + 2}$, and $\hat{\mathbf{j}}$ is a unit vector normal to the dielectric surface.

By substituting Equations \ref{dep_force}, and \ref{image_force} into \ref{force_density} we have
\begin{eqnarray*}
 \mathbf{F}_E &=& q \mathbf{E} + \mathbf{F}_{DEP} + \mathbf{F}_I \\
 &=& q \mathbf{E} + \frac{k q^2}{16 \pi \epsilon_0 } y^{-2} \hat{\mathbf{j}} + 2 \pi R_d^3 \kappa_w \epsilon_0 K \nabla E^2, 
\end{eqnarray*}
and the 1-D governing equation becomes
\begin{equation}
 \label{gov_eqn_subs}
m y'' = - \frac{1}{2} C_D \rho A {y'}^2 - q E - \frac{k q^2}{16 \pi \epsilon_0} y^{-2}- 2 \pi R_d^3 \kappa_w \epsilon_0 K \nabla E^2,
\end{equation}
subject to
\begin{equation}
y(0) = R, \hspace{5 mm} \mbox{and} \hspace{5 mm} y'(0) = U_0 .
\end{equation}

By comparing DEP and Coulombic terms in Equation \ref{gov_eqn_subs}, we note that a condition to neglect the DEP force is
\begin{eqnarray}
\frac{ \kappa_w \epsilon_0 K R_d^2 E_0}{q} \ll 1. \nonumber
\end{eqnarray}
As this condition likely prevails in our experiments we henceforth neglect the DEP force. There is some physical intuition to support this conclusion as well. The dielectric displacement $\mathbf{D} = \kappa \epsilon_0 \mathbf{E}$ of a water drop in air is high due to its large relative dielectric constant. This implies that the field strength within the drop is about 80 times smaller than in the surrounding medium, which is essentially the same as a vacuum from a dielectric standpoint. Thus it is not particularly inaccurate to model the dielectric volume of a drop as an equipotential conductive shell with zero field in its interior. As an aside, in treating the drop as an ideal conductor we note that in our specific case the electrostatic force is not a body force \emph{per se} as the electric field is acting on charges on the surface of the conductor.
